---
title: "Explore and Mine Data"
author: "Pratyush Rokade"
date: "04/16/2024"
output: pdf_document
---

```{r importLibraries, echo=FALSE}
pacman::p_load(RMySQL, ggplot2)
```

## Report

```{r connectDatabase, echo=FALSE}
db <- dbConnect(RMySQL::MySQL(),
                dbname = "sql5699558",
                host = "sql5.freemysqlhosting.net",
                user = "sql5699558",
                password = "Ip2dMxppa6")
```

## Analytical Query 1

```{r query1, echo=FALSE}

# Retrieve data on the top sales reps for each year
analytical_query1 <- dbGetQuery(db, "
SELECT name, year, total_sold
FROM (
    SELECT 
        r1.year,
        r1.name,
        SUM(r1.total_sold) AS total_sold,
        (
            SELECT COUNT(DISTINCT r2.total_sold) + 1
            FROM rep_facts AS r2
            WHERE r2.year = r1.year AND r2.total_sold > SUM(r1.total_sold)
        ) AS rank
    FROM rep_facts AS r1
    GROUP BY r1.year, r1.name
) AS ranked_sales
WHERE rank = 1
ORDER BY year, total_sold DESC;
")

analytical_query1


print(endt-startt)

# Plotting the data
ggplot(analytical_query1, aes(x = factor(year), y = total_sold, fill = name)) +
  geom_bar(stat = "identity", position = "dodge", colour = "black") +
  labs(title = "Top 5 Sales Reps with the most sales broken down by Year",
       x = "Year",
       y = "Total Amount Sold",
       fill = "Name") +
  scale_fill_manual(values = c("red", "blue", "green", "yellow", "orange")) +
  theme(plot.title = element_text(hjust = 0.5))
```
From the above plot it is clearly visible the sales figures of the top 5 reps each year. Walison da Silva has made the highest sales in each of the three years, whereas Helmut Schwab has made the least sales across the three years.


## Analytical Query 2

```{r query2, echo=FALSE}
# Retrieve quarterly sales data
analytical_query2 <- dbGetQuery(db, "
SELECT 
    name,
    year,
    quarter,
    SUM(total_amount_sold) as total_amount_sold
FROM 
    product_facts
GROUP BY 
    name,
    year,
    quarter
ORDER BY 
    name,
    year,
    quarter;
")

# Create a new column combining year and quarter for better visualization
analytical_query2$time <- with(analytical_query2,
                               paste(year, quarter, sep=" Q"))

# Plotting the quarterly sales data
ggplot(data = analytical_query2, aes(x = time, y = total_amount_sold,
                                     color = name, group = name)) +
  geom_line() + 
  geom_point() + 
  labs(title = "Quarterly Sales Data",
       x = "Year - Quarter",
       y = "Total Amount Sold",
       caption = "Data: Company Sales") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5))

analytical_query2
```
Observing the above plot, Zalofen has had the highest sales made in all of the quarters from 2020 through 2022. Diaprogenix has made the lowest sales across all the quarters in the three years.


## Analytical Query 3

```{r query3, echo=FALSE}
# Retrieve total units sold per product per country
analytical_query3 <- dbGetQuery(db, "
SELECT 
    name,
    country,
    SUM(units_sold) AS total_units_sold
FROM 
    product_facts 
GROUP BY 
    name,
    country;
")

# Plotting total units sold per country with separate lines for each product
ggplot(analytical_query3, aes(x = country, y = total_units_sold,
                              group = name, color = name)) +
  geom_line() +
  geom_point() +
  labs(title = "Total Units Sold Per Country",
       x = "Country",
       y = "Total Units Sold",
       color = "Product") +
  theme(plot.title = element_text(hjust = 0.5))

# Plotting total units sold per country with separate facets for each product
ggplot(analytical_query3, aes(x = country, y = total_units_sold,
                              group = name)) +
  geom_line() +
  geom_point() +
  labs(title = "Total Units Sold Per Country",
       x = "Country",
       y = "Total Units Sold") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 0.75)) +
  facet_wrap(~name, scales = "free_y")
```
The first plot shows the comparison of the number of units being sold for all products across the three countries. The second plot shows individual line charts of total units sold for each product across the three countries.


## Analytical Query 4

```{r query4, echo=FALSE}
# Retrieve average annual sales per sales rep over the years
analytical_query4 <- dbGetQuery(db, "
SELECT 
    rep_id,
    name,
    year,
    AVG(total_sold) AS average_annual_sales
FROM 
    rep_facts
GROUP BY 
    rep_id, 
    year
ORDER BY
    name;
")

# Plotting average sales per sales reps over the years
ggplot(data = analytical_query4, aes(x = year, y = average_annual_sales,
                                     group = name, color = name)) +
  geom_line() +  
  geom_point() +  
  labs(title = "Average Sales per Sales Reps over the Years",
       x = "Year",
       y = "Average Annual Sales",
       color = "Sales Rep") +
  scale_x_continuous(breaks = unique(analytical_query4$year)) +
  theme(plot.title = element_text(hjust = 0.5))  

```
Walison da Silva has had the highest average sales across all the years, whereas Veronica Sixt has achieved the least average sales.

```{r disconnectDatabase, echo=FALSE}
dbDisconnect(db)
```